Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-06-26 doc/manuscript_hkk_drought_sensitivity.Rmd
abstract suggestions

# Abstract (150 words) {-}
Increased frequency and severity of droughts threatens forest health worldwide, with important implications for this critical carbon sink. 
Tree species adaptations to hot and dry conditions--e.g., dry-season deciduousness in tropical seasonal forests--might confer drought resistance, but such resistance may be variable across species, microenvironments, and drought events. 
We analysed dendrometer band measurements on 1820 trees across 30 species in a seasonally dry tropical forest in Western Thailand to assess tree growth responses to three climatically distinct droughts. 
Species and individuals exhibited a wide range of growth responses within each drought, and differences in response intensity and affect among the drought events. 
Deciduous and evergreen species were more sensitive to wet- and dry-season drought respectively, and wetter microclimates amplified this difference. 
<!-- Exposed trees showed more sensitivity to drought, but  larger sized trees or trees in wetter topographic positions seemed to buffer growth declines.  -->
Heterogeneous drought responses of species and individuals indicate potential spatiotemporal insurance effects in diverse forests in the face of increased drought.

>>>>

# Abstract (150 words) {-}
Increased frequency and severity of droughts threatens forest health worldwide, with important implications for this critical carbon sink. 
Tree species adaptations to hot and dry conditions -- e.g., dry-season deciduousness in tropical seasonal forests -- can confer drought resistance, but such resistance may be variable across species, microenvironments, and drought events. 
Here, we assess growth responses of 1820 trees across 30 species to three climatically distinct droughts in a seasonally dry tropical forest in Western Thailand. 
Species and individuals exhibited a wide range of growth responses within each drought, and differences in response intensity and affect among the drought events. 
Deciduous and evergreen species were more sensitive to wet- and dry-season drought respectively<!--, and wetter microclimates amplified this difference-->.
While crown exposure consistently increased drought sensitivity, topographic wetness had variable effects.
<!-- Exposed trees showed more sensitivity to drought, but  larger sized trees or trees in wetter topographic positions seemed to buffer growth declines.  -->
Heterogeneous drought responses of species and individuals indicate potential spatiotemporal insurance effects in diverse forests in the face of increased drought.
